Forecasting the 2024 U.S. Presidential Election through Poll Aggregation and Adjustments for Poll Quality

### 2024-11-04 paper/paper.qmd
update paper table and graph
made text understandability edits

The objective of this study is to predict the support percentages for the primary candidates, Kamala Harris and Donald Trump, using this hybrid model. We incorporate aggregation by filtering the polls on @fivethirtyeight by numeric grade that indicates pollster’s reliability. We also quantify other polling attributes, such as sample size, poll reliability scores, and polling methodology, which serve as predictors. Our results show that this model provides a more stable and accurate forecast than traditional poll aggregation alone.

The remainder of this paper is structured as follows: @sec-data provides an overview of the data. @sec-model provides the modeling approach, including simple and multiple linear regression models and a Bayesian hierarchical model. We then present our results in @sec-result and discuss the implications, limitations, and future research directions in @sec-discussion.

@@

The estimand for this research paper is the predicted support percentages for Kamala Harris and Donald Trump. The prediction is based on quantifying various polling factors, including sample size, poll scores, and transparency scores, which are used as predictors.

@@

- pct: The percentage of support or vote share that each candidate (Kamala Harris or Donald Trump) received in the poll.

@@

- state: The U.S. state in which the poll was conducted or targeted.
- methodology: The method used to conduct the poll (e.g., online, phone, in-person).
- party: The political party of the candidate (e.g., DEM for Democrat, REP for Republican).
- candidate_name: The name of the candidate, either Kamala Harris or Donald Trump.
- pollster: The name of the polling organization that conducted the poll.
- stage: The stage of the election (e.g., "general").

@@

We then create a new attribute, days_taken_from_election, which represents the time gap between the poll's start date and the actual U.S. election date. Additionally, we filter the dataset to include only polls conducted after July 21, 2024, the date when Kamala Harris declared her candidacy. Finally, we remove any remaining rows that contain missing values to ensure a clean dataset.

@@

This absence of polling data is not a significant concern due to the structure of the United States Electoral College (explained in detail in the Appendix). The states lacking polling data have consistently followed historical voting patterns, so predicting the winning candidate in those states is unnecessary.

@@

### Basic Statistics Summary for Cleaned Data

@@

cleaned_data %>%
  filter(candidate_name %in% c("Kamala Harris", "Donald Trump")) %>%
  group_by(state, candidate_name) %>%
  summarise(avg_pct = mean(pct, na.rm = TRUE), .groups = "drop") %>%
  ggplot(mapping = aes(x = avg_pct, y = state, fill = candidate_name)) + 
  geom_col(position = "stack") +
  theme_minimal() +
  labs(
    x = "Average Polled Support (%)", 
    y = "State",
    fill = "Candidate"
  ) +
  scale_fill_manual(values = c("Kamala Harris" = light_blue, "Donald Trump" = light_red)) +
  theme_minimal() +
  theme(legend.position = "bottom") + 
  theme(axis.text.y = element_text(size = 8))

@@

In @fig-avg-pct-state, historically Democratic are polling for Kamala and historically Republican states are polling for Trump. Similarly, historically swing states also appear to be close, for instance Michigan (46.9% Trump, 47.5% Harris), Nevada (47.7% Trump, 47.9% Harris), and Pennsylvania (46.9% Trump, 48.2% Harris) all are close to an even split.

@@

@fig-avg-pct-methodology shows that polls utilizing multiple communication methods to reach voters tend to have lower interquartile ranges (IQRs) in their boxplots compared to those that rely on only one or two communication methods.

```{r}
#| label: fig-pollscore-samplesize
#| fig-cap: More Reliable Pollsters Have Larger Sample Sizes in their Polls
#| echo: false
#| warning: false

@@

@@

Based on @fig-pollsamplesize-canidate-pct, there does not seem to be a relationship between the sample size of a poll the percentage of a candidate.

@@

@@

We chose sample size, poll score, and days taken from the election as predictors for building the model based on the visual trends observed in the plots. As shown in @fig-pollsamplesize-canidate-pct, @fig-pollscore-candidate-pct and @fig-polldate-canidate-pct in @sec-data, the relationship between pct and sample size shows that larger sample sizes correspond to more stable percentages, indicating that sample size could significantly impact prediction accuracy. The plot of pct versus poll score shows a negative trend, suggesting that poll scores are consistently influential in explaining variations in pct. Additionally, the pct versus days taken plot indicates that the proximity to the election affects the percentage, with a slight negative trend showing that time remaining has predictive value. These observed patterns imply that these variables can effectively capture key factors influencing pct and are suitable for model inclusion.

@@

@@

# Trump Voter Prediction Model

The multiple linear regression model (MLR) for Donald Trump will use the same variables, formula, and Bayesian approach as the one for Harris. Likewise, the Trump dataset is also split into training and testing data. The model results are similar to that of Harris's MLR model.

Based on @fig-model-trumpmlr-res, the test data predictions are also close to the actual values. This suggests that model can generalize outside data. Similarly, the distance between the prediction and the actual values do not appear to follow a pattern, suggesting that the error is due to randomness and not model bias.

# Trump Model Results

```{r}
#| label: fig-model-trumpmlr
#| fig-cap: "MLR Trump Model Accounts For A Large Amount of Variability in Voter Percentage"
#| echo: false
#| warning: false
just_trump <- cleaned_data %>% filter(candidate_name == "Donald Trump")
predictions2 = posterior_predict(bayesian_model_train_trump, newdata = just_trump)
predicted_means = colMeans(predictions2)
predicted_intervals = apply(predictions2, 2, quantile, probs = c(0.025, 0.975))
result_summary = data.frame(
  Actual_PCT = just_trump$pct,
  Predicted_PCT = predicted_means,
  Lower_CI = predicted_intervals[1, ],
  Upper_CI = predicted_intervals[2, ]
)
ggplot(result_summary, aes(x = result_summary$Actual_PCT, y = result_summary$Predicted_PCT)) +
  geom_point(alpha = 0.7) +
  geom_errorbar(aes(ymin = Lower_CI, ymax = Upper_CI), width = 0.2, color = "blue") +
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
  labs(
    title = "Actual vs Predicted Percentage For Trump MLR Model",
    x = "Actual PCT",
    y = "Predicted PCT"
  )
```

@@

## Election Prediction {#sec-prediction}

Our prediction process consists of two primary components. First, we develop models for both Trump and Harris based on the variables outlined in @sec-model. This involves partitioning the dataset into training and testing subsets. Next, we further divide the testing dataset into swing states and other competitive races. We then input this test data into the respective models to generate predictions. By averaging these predictions, we can calculate the expected voter percentage for each candidate in each state. The candidate with a higher percentage is deemed the winner for that state.

We generated predictions for the following states: Arizona, Nevada, Georgia, Pennsylvania, Michigan, Minnesota, Wisconsin, Florida, Texas, Maine CD-2, Nebraska CD-2, New Hampshire, Ohio, Virginia, North Carolina, and Iowa. Winners for other states were determined based on historical trends and predictions from sources like @cnn_electoral_nodate. Most states without predictions are strongly Republican or Democratic, so their absence is not expected to significantly impact prediction validity.

@@

Based on the figures in @sec-data, the team identified three sources of bias beyond sample size and state: methodology, the timing of polls relative to the election date, and poll scores. From the analysis in Figure @fig-avg-pct-methodology, we found that polls employing a limited number of methods to reach and communicate with voters exhibited greater variability. This finding is plausible, as different demographic groups often prefer distinct communication methods; for example, older voters may favor phone surveys, while younger voters might prefer online surveys. Additionally, utilizing multiple communication channels provides pollsters with more opportunities to contact potential voters, thereby reducing non-response bias. Consequently, it is essential to incorporate methodology as a variable in the model, and we recommend that pollsters adopt diverse communication methods whenever possible.

From @fig-polldate-canidate-pct, the team observed a decline in third-party candidate support as the election approached. Early in the election cycle, voters may entertain the idea of a viable third-party candidate, but as the election nears, interest in third-party options diminishes. This trend may be worsened by decreased participation in polls as the election date approaches. As a result, the people who do participate in those earlier polls are likely to be more informed and passionate, which are qualities that may be associated with support for third-party candidates. Additionally, the withdrawal of a prominent third-party candidate, Robert F. Kennedy Jr., on August 23, 2024 (74 days from the election), could influence this dynamic. However, we find the earlier reasons more compelling, as the data in Figure @fig-polldate-canidate-pct do not indicate a significant drop in third-party support at that time, but rather a gradual decline.

Finally, as illustrated in Figure @fig-pollscore-candidate-pct, more reliable polls tend to show a bias favoring Trump over Harris—not necessarily indicating a guaranteed victory for Trump, but rather suggesting he may outperform expectations in many polls. It is important to note that the FiveThirtyEight poll score system is partly based on historical accuracy. Given Trump's overperformance in the polls during the 2016 and 2020 elections, this scoring may be overcorrecting for past inaccuracies. This potential undervaluation of Trump’s support could be strategically beneficial for the Harris campaign, providing a warning to avoid repeating previous electoral mistakes.

@@

@@

In @sec-prediction, the team predicted that Kamala Harris would secure 298 delegates while Donald Trump would obtain 240 delegates. However, this prediction has several limitations. Firstly, the model used to generate these predictions inherits the constraints outlined in @sec-model-limitations. Additionally, the predictions are based on averaging results from polling data for each state, which can lead to propagated prediction errors.

Each state has a limited number of polls available; while the overall analysis considered a substantial number of polls, individual states often had fewer than 20 polls, with even the largest states contributing only up to 27. Moreover, as noted in Section @sec-prediction, the analysis focused solely on contested races, specifically swing states. Consequently, many states were assigned to a candidate based on strong historical trends rather than direct predictions. Although this approach is reasonable, it may overlook emerging trends that could potentially shift a state's allegiance from Republican to Democrat or vice versa.

@@

The American federal or national government is split into three branches, executive., legislative, and judicial the executive branch includes the president and the military. The legislative branch includes two subgroups, the House of Representatives, and the Senate. These two subgroups create laws. Every state has two Senators and one Representative per approximately 750,000 people. The judicial branch is the court system.

@@

The Electoral College is the system used in the United States to elect the president and vice president. Instead of a direct popular vote, each state is allocated a certain number of electors based on its representation in Congress (the total of its Senators and Representatives). When voters vote, they are voting for a slate of electors pledged to a candidate. The candidate who receives the most electoral votes (270 out of 538) wins the presidency. This system means that winning the popular vote in a state result in winning all that state's electoral votes. The only exception are the states of Maine and Nebraska, who award electoral votes by congressional district, with two additional votes given to the statewide winner.

The Electoral College results in some states being unnecessary to campaign in, as their strong historical voting pasterns towards either Democrats or Republicans make them unlikely to change, regardless of campaign efforts. Therefore, for statisticians, polling information from these states may not be that useful when trying to predict the outcome of an election. On the other hand, states that can vote either Democratic or Republican (swing states) are immensely important when predicting an election. As a result, campaigns spend hundreds of millions of dollars campaigning and understanding voters there. 
```{r}
#| label: fig-currentstateofelection
#| fig-cap: 2024 U.S. Presidential Election State Forecast Map
#| echo: false
#| warning: false
light_blue <- "#ADD8E6"
dark_blue <- "#6688CC"
light_red <- "#FF7777"
dark_red <- "#B22222"
purple <- "#9370DB"
# Define your colors for each state (replace these colors with your desired ones)
state_colors_legend <- c(
  "Republican" = dark_red,      # Dark Red
  "Lean Republican" = light_red, # Light Red
  "Toss-Up" = purple,         # Purple
  "Lean Democrat" = light_blue,   # Light Blue
  "Democrat" = dark_blue,        # Dark Blue
  "Split By District" = "gray"
)

@@

Notably, Nebraska and Maine are indicated in gray due to their delegates being split by district. In Nebraska, the state overall is projected to lean Republican, with the first and third districts also strongly favoring Republican candidates, while the second district leans Democratic. In Maine, the overall expectation is a Democratic leaning, consistent with its first district, though the second district leans Republican. Furthermore, Alaska and Hawaii are not on the map. Alaska is strongly favoring Republicans while Hawaii strongly favors democrats.

The seven agreed upon swing states are Pennsylvania, North Carolina, Georgia, Arizona, Nevada, Wisconsin, and Michigan with Texas, Florida, Nebraska District Two, Maine District 2, and Minnesota as the next closest races.

@@

@@

Given the monetary constraint, the team's first decision is to mostly include swing states and districts. These are Nevada, Arizona, Wisconsin, Michigan, Pennsylvania, North Carolina, and Georgia. Swing states are the primary focus of the poll because they disproportionately affect the outcome of the election based on the explanation in @sec-electoral-college. However, the poll will also include Texas, Florida, Minnesota, Nebraska District 2, and Maine District 2, as they have the potential to one by either party but are not as likely to change as the previously mentioned swing states. Nevertheless, it is important to note that to save costs, the team will focus on polling true swing states compared to Texas, Florida, and Minnesota.

@@

The team will then find and use multiple databases, such as @usps_customer_nodate, to find addresses, telephones, and emails of potential voters in those states. According to @pew_research_center_us_nodate, gathering this information reduces non-response and selection bias, as the team can contact the same individual through multiple mediums. Moreover, certain demographics may prefer specific communication forms; for instance, the elderly may prefer phone or paper mail polls over text or email. Additionally, to encourage individuals to complete the survey, one in every 50 participants will have a chance at winning $20.

@@

Each state will have its own poll, but the polling methodology and the survey given will remain the same. Specifically, the polls will use stratified random sampling to make sure that participants reflect each state's counties in proportion to their population sizes. County proportion is used as the stratification variable because it is easily accessible. Additionally, counties possess inherent demographic and cultural characteristics that can influence their alignment with particular political parties. Ideally, stratification would also ensure race, gender, age, and other socioeconomic factors are also accounted for. However, these factors are almost impossible to determine while sending the survey. Therefore, the research team decided to ask demographic questions directly inside the survey. After data collection, the team will weigh data based on demographics. For example, if a certain county has a 30% black population (this statistic will be determined from the US census), but only 15% of the survey participants are black, then the pollsters may decide to count each black participant's responses twice.

Furthermore, the team will aim to sample approximately 1,000 people from each true swing state. This sample size is large enough to provide reliable conclusions but not so large that it resembles sampling with replacement. For true swing states, the team aims to survey 1,000 individuals. However, if the final sample size falls short of this target, it is not considered a significant issue.

@@

The team's ideal survey will be made using @qualtrics_qualtrics_nodate. It will attempt to ensure four things: no leading questions, no question order bias, no answer order bias, and the survey should identify non-engaged participants. To identify participants who are not engaged, the research team has decided to add a worthless question. This question is extremely simple and clearly has one correct answer. Therefore, if a participant selects the wrong answer, they are not engaged with the survey and their responses should be removed from the final data. An example of this is question number 7.

@@

Answer bias, like order bias, occurs when the order of answer choices influences participants' selections, with the first option often chosen more frequently. To prevent this, answer choices should be randomized upon survey entry. Questions 8 & 9 could benefit from this. Likewise, the team's hypothetical survey has not implemented this feature, but it is available in the @qualtrics_qualtrics_nodate paid plan.

@@

While the team's methodology and survey create a robust system, there are potential issues. The reliance of weighting results based on a candidate's demographics can lead to error propagation. For instance, if a certain racial demographic population is only captured limitedly and that limited sample is far from representative, a few individuals in the population can have impact on the polls prediction of what candidate will win the state. There is also a selection bias in terms of the monetary reward. Potentially, people who like monetary rewards would be more likely to engage in the survey and could therefore exhibit certain voting or demographic characteristics that create an unrepresentative sample.

>>>>

The objective of this study is to predict the support percentages for the primary candidates, Kamala Harris and Donald Trump, using this hybrid model. We incorporate aggregation by filtering the polls on @fivethirtyeight by numeric grade that indicates pollster’s reliability. We also quantify other polling attributes, such as sample size, poll reliability scores, and polling methodology, which serve as predictors. Our results show that this model provides a more stable and accurate forecast than traditional poll aggregation alone. With this model, we generate a prediction that Kamala Harris will win the popular vote in 2024 and win the electoral college.

The remainder of this paper is structured as follows: @sec-data provides an overview of the data. @sec-model provides the modeling approach, including simple and multiple linear regression models and a Bayesian hierarchical model. We then present our results in @sec-result and discuss the implications, limitations, then we predict the election in @sec-prediction, and future research directions in @sec-discussion.

@@

The estimand for this research paper is the predicted support percentages for Kamala Harris and Donald Trump. The prediction is based on quantifying various polling factors, including sample size, poll scores, and poll methodology, which are used as predictors.

@@

- pct: The percentage of support or vote share that each candidate (Kamala Harris, Donald Trump, or third party alternatives) received in the poll.

@@

- state: The U.S. state in which the poll was conducted or targeted.
- methodology: The method used to conduct the poll (e.g., online, phone, in-person).
- party: The political party of the candidate (e.g., DEM for Democrat, REP for Republican).
- candidate_name: The name of the candidate (i.e. Kamala Harris, Donald Trump, or third party candidates).
- pollster: The name of the polling organization that conducted the poll.
- stage: The stage of the election (e.g., "general").

@@

We then create a new attribute, days taken from election, which represents the time gap between the poll's start date and the actual U.S. election date. Additionally, we filter the dataset to include only polls conducted after July 21, 2024, the date when Kamala Harris declared her candidacy. Finally, we remove any remaining rows that contain missing values to ensure a clean dataset.

@@

This absence of polling data is not a significant concern due to the structure of the United States Electoral College (explained in detail in @sec-electoral-college). The states lacking polling data have consistently followed historical voting patterns, so predicting the winning candidate in those states is unnecessary.

@@

#### State and Candidate Percentage

@@

cleaned_data %>%
  filter(candidate_name %in% c("Kamala Harris", "Donald Trump")) %>%
  group_by(state, candidate_name) %>%
  summarise(avg_pct = mean(pct, na.rm = TRUE), .groups = "drop") %>%
  ggplot(mapping = aes(x = avg_pct, y = state, fill = candidate_name)) + 
  geom_col(position = "stack") +
  theme_minimal() +
  labs(
    x = "Average Polled Support (%)", 
    y = "State",
    fill = "Candidate"
  ) +
  scale_fill_manual(values = c("Kamala Harris" = light_blue, "Donald Trump" = light_red)) +
  theme_minimal() +
  theme(legend.position = "bottom") + 
  theme(axis.text.y = element_text(size = 5))

@@

In @fig-avg-pct-state, historically Democratic states are polling for Kamala and historically Republican states are polling for Trump. Similarly, historically swing states also appear to be close, for instance Michigan (46.9% Trump, 47.5% Harris), Nevada (47.7% Trump, 47.9% Harris), and Pennsylvania (46.9% Trump, 48.2% Harris) all are close to an even split.

#### Poll Methodology and Candidate Percentage

@@

@fig-avg-pct-methodology shows that polls utilizing multiple communication methods to reach voters tend to have lower interquartile ranges (IQRs) in their boxplots compared to those that rely on only one or two communication methods. It is important to note that statistical outliers are not visible in the graph to ensure better visibility.

#### Poll Score and Sample Size

```{r}
#| label: fig-pollscore-samplesize
#| fig-cap: Pollsters with Higher Pollscores Have Larger Sample Sizes in their Polls
#| echo: false
#| warning: false

@@

#### Sample Size and Voter Percentage

@@

Based on @fig-pollsamplesize-canidate-pct, there does not seem to be a relationship between the sample size of a poll the voter percentage of a candidate.

#### Poll Score and Voter Percentage

@@

#### Days from Election and Voter Percentage

@@

We chose sample size, poll score, and days taken from the election as predictors for building the model based on the visual trends observed in the plots. As shown in @fig-pollsamplesize-canidate-pct, @fig-pollscore-candidate-pct and @fig-polldate-canidate-pct in @sec-data, the relationship between percentage and sample size shows that larger sample sizes correspond to more stable percentages, indicating that sample size could significantly impact prediction accuracy. The plot of percentage versus poll score shows a negative trend for Donald Trump and a positive trend for Kamala Harris, suggesting that poll scores are consistently influential in explaining variations in percentage. Additionally, the percentage versus days taken plot indicates that the proximity to the election affects the percentage, with a slight boost towards major candidates. These observed patterns imply that these variables can effectively capture key factors influencing pct and are suitable for model inclusion.

@@

The team also built a Bayesian MLR using the same approach and formula for Trump. It is in section @sec-trumpmodel.

@@

The team has provided results for a Trump Bayesian MLR model. It is in section @sec-trumpmodelresults.

@@

# Election Prediction {#sec-prediction}

Our prediction process consists of two primary components. First, we develop models for both Trump and Harris based on the variables outlined in @sec-model and @sec-trumpmodel. This involves partitioning the dataset into training and testing subsets. Next, we further divide the testing dataset by state. We then input this test data into the respective models to generate predictions. By averaging these predictions, we can calculate the expected voter percentage for each candidate in each state. The candidate with a higher percentage is deemed the winner for that state.

We generated predictions for the following states: Arizona, Nevada, Georgia, Pennsylvania, Michigan, Minnesota, Wisconsin, Florida, Texas, Maine CD-2, Nebraska CD-2, New Hampshire, Ohio, Virginia, North Carolina, and Iowa. Winners for other states were determined based on historical trends and predictions from sources like @cnn_electoral_nodate. Most states without predictions are strongly Republican or Democratic, so their absence is not expected to significantly impact prediction validity.
```{r}
#| echo: false
#| warning: false
set.seed(123)
train_data_trump <- read_parquet(here("data/02-analysis_data/train_data_trump.parquet"))
test_data_trump <- read_parquet(here("data/02-analysis_data/test_data_trump.parquet"))
bayesian_model_train_trump <- readRDS(here("models/bayesian_model_train_trump.rds"))
```

@@

Based on the figures in @sec-data, the team identified three sources of bias beyond sample size and state: methodology, the timing of polls relative to the election date, and poll scores. From the analysis in Figure @fig-avg-pct-methodology, we found that polls employing a limited number of methods to reach and communicate with voters exhibited greater variability. This finding is plausible, as different demographic groups often prefer distinct communication methods; for example, older voters may favor phone surveys, while younger voters might prefer online surveys. Additionally, using multiple communication methods provides pollsters with more opportunities to contact potential voters, thereby reducing non-response bias. Consequently, it is essential to incorporate methodology as a variable in the model, and we recommend that pollsters adopt diverse communication methods whenever possible.

From @fig-polldate-canidate-pct, the team observed a decline in third-party candidate support as the election approached. Early in the election cycle, voters may entertain the idea of a viable third-party candidate, but as the election nears, interest in third-party options diminishes. This trend may be worsened by decreased participation in polls further from the election day. As a result, the people who do participate in those earlier polls are likely to be more informed and passionate, which are qualities that may be associated with support for third-party candidates. Additionally, the withdrawal of a prominent third-party candidate, Robert F. Kennedy Jr., on August 23, 2024 (74 days from the election), could influence this dynamic. However, we find the earlier reasons more compelling, as the data in Figure @fig-polldate-canidate-pct do not indicate a significant drop in third-party support at that time, but rather a gradual decline.

Finally, as illustrated in Figure @fig-pollscore-candidate-pct, more reliable polls tend to show a bias favoring Trump over Harris—not necessarily indicating a guaranteed victory for Trump, but rather suggesting he may outperform expectations in many polls. It is important to note that the FiveThirtyEight poll score system is partly based on historical accuracy. Given Trump's overperformance in the polls during the 2016 and 2020 elections, this scoring may be overcorrecting for past inaccuracies. This potential undervaluation of Trump’s support could benefit the Harris campaign, by providing a warning to not repeat previous election mistakes.

@@

Furthermore, the incorporation of both poll score and sample size into the model may lead to the potential for double counting the sample size. This concern was raised in Nate Silver's article [@silver_polls_2008]. As illustrated in @fig-pollscore-samplesize, there is a positive correlation between poll score and sample size. However, this correlation is weak. Additionally, given that pollsters with negative poll scores are generally more reliable than those with positive scores, we would expect the opposite relationship to hold true for sample size to even have a possibility of being double counted.

@@

In @sec-prediction, the team predicted that Kamala Harris would win 298 delegates while Donald Trump would get 240 delegates. However, this prediction has several limitations. First, the prediction will inherit all of the limitations of the model it (model limitations in @sec-model-limitations). Additionally, the predictions are based on averaging results from polling data for each state, which can lead to propagated prediction errors.

Also, each state has a limited number of polls available. While the overall analysis considered a substantial number of polls, individual states often had fewer than 20 polls, with even the largest states having only 27. Moreover, as noted in section @sec-prediction, the analysis focused solely on contested races, specifically swing states. Therefore, many states were assigned to a candidate based on strong historical trends and not from predictions. Although this approach is reasonable, it may overlook current trends that could potentially shift a state's vote from Republican to Democrat or vice versa.

@@

The American federal or national government is split into three branches, executive, legislative, and judicial. The executive branch includes the president and the military. The legislative branch includes two subgroups, the House of Representatives, and the Senate. These two subgroups create laws. Every state has two Senators and one Representative per approximately 750,000 people. Finally, the judicial branch is the court system.

@@

The Electoral College is the system used in the United States to elect the president and vice president. Instead of a direct popular vote, each state is allocated a certain number of electors based on its representation in Congress (the total of its Senators and Representatives). When voters vote, they are voting for what candidate they want their electors to vote for. The candidate who receives the most electoral votes (270 out of 538) wins the presidency. This system means that winning the popular vote in a state results in winning all that state's electoral votes. The only exception are the states of Maine and Nebraska, who award electoral votes by congressional district, with two additional votes given to the statewide winner.

The Electoral College results in some states being unnecessary to campaign in, as their strong historical voting patterns towards either Democrats or Republicans make them unlikely to change, regardless of campaign efforts. Therefore, for statisticians, polling information from these states may not be that useful when trying to predict the outcome of an election. On the other hand, states that can vote either Democratic or Republican (swing states) are immensely important when predicting an election. As a result, campaigns spend hundreds of millions of dollars campaigning and understanding voters there. 
```{r}
#| label: fig-currentstateofelection
#| fig-cap: 2024 U.S. Presidential Election State Forecast Map
#| echo: false
#| warning: false
light_blue <- "#ADD8E6"
dark_blue <- "#6688CC"
light_red <- "#FF7777"
dark_red <- "#B22222"
purple <- "#9370DB"
# Define your colors for each state (replace these colors with your desired ones)
state_colors_legend <- c(
  "Republican" = dark_red,      # Dark Red
  "Lean Republican" = light_red, # Light Red
  "Toss-Up" = purple,         # Purple
  "Lean Democrat" = light_blue,   # Light Blue
  "Democrat" = dark_blue,        # Dark Blue
  "Split By District" = "gray"
)

@@

Notably, Nebraska and Maine are indicated in gray due to their delegates being split by district. In Nebraska, the state overall is projected to lean Republican, with the first and third districts also strongly favoring Republican candidates, while the second district leans Democratic. In Maine, the overall state is a Democratic leaning, and so is its first district, but the second district leans Republican. Furthermore, Alaska and Hawaii are not on the map. Alaska is strongly favoring Republicans while Hawaii strongly favors democrats.

The seven agreed upon swing states are Pennsylvania, North Carolina, Georgia, Arizona, Nevada, Wisconsin, and Michigan with Texas, Florida, Nebraska District Two, Maine District 2, Iowa, and Minnesota as the next closest races.

@@

# Trump Voter Prediction Model {#sec-trumpmodel}

The multiple linear regression model (MLR) Bayesian hierarchical model for Donald Trump will use the same variables, formula, and Bayesian approach as the one for Harris. Likewise, the Trump dataset is also split into training and testing data. The model results are similar to that of Harris's MLR model.

Based on @fig-model-trumpmlr-res, the test data predictions are also close to the actual values. This suggests that model has predictability strength. Similarly, the distance between the prediction and the actual values do not appear to follow a pattern, suggesting that the error is due to randomness and not model bias.

# Trump Model Results {#sec-trumpmodelresults}

```{r}
#| label: fig-model-trumpmlr
#| fig-cap: "Bayesian MLR Trump Model Accounts For A Large Amount of Variability in Voter Percentage"
#| echo: false
#| warning: false
just_trump <- cleaned_data %>% filter(candidate_name == "Donald Trump")
predictions2 = posterior_predict(bayesian_model_train_trump, newdata = just_trump)
predicted_means = colMeans(predictions2)
predicted_intervals = apply(predictions2, 2, quantile, probs = c(0.025, 0.975))
result_summary = data.frame(
  Actual_PCT = just_trump$pct,
  Predicted_PCT = predicted_means,
  Lower_CI = predicted_intervals[1, ],
  Upper_CI = predicted_intervals[2, ]
)
ggplot(result_summary, aes(x = result_summary$Actual_PCT, y = result_summary$Predicted_PCT)) +
  geom_point(alpha = 0.7) +
  geom_errorbar(aes(ymin = Lower_CI, ymax = Upper_CI), width = 0.2, color = "blue") +
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
  labs(
    title = "Actual vs Predicted Percentage For Trump MLR Model",
    x = "Actual PCT",
    y = "Predicted PCT"
  )
```

@@

Given the monetary constraint, the team's first decision is to mostly include swing states and districts. These are Nevada, Arizona, Wisconsin, Michigan, Pennsylvania, North Carolina, Georgia, Nebraska District 2, and Maine District 2. Swing states are the primary focus of the poll because they disproportionately affect the outcome of the election based on the explanation in @sec-electoral-college. However, the poll will also include other close races like Texas, Florida, and Minnesota, as they have the potential to be won by either party but are not as likely to change as the previously mentioned swing states. Nevertheless, it is important to note that to save costs, the team will focus on polling true swing states compared to states like Texas, Florida, and Minnesota.

@@

The team will then find and use multiple databases, such as U.S. Postal Service’s residential address file [@usps_customer_nodate], to find addresses, telephones, and emails of potential voters in those states. According to @pew_research_center_us_nodate, gathering this information reduces non-response and selection bias, as the team can contact the same individual through multiple mediums. Moreover, certain demographics may prefer specific communication forms; for instance, the elderly may prefer phone or paper mail polls over text or email. Additionally, to encourage survey completion, one in every 50 participants will have a chance at winning $20.

@@

Each state will have its own poll, but the polling methodology and the survey given will remain the same. Specifically, the polls will use stratified random sampling to make sure that participants reflect each state's counties in proportion to their population sizes. County proportion is used as the stratification variable because that information is easily accessible. Additionally, counties possess inherent demographic and cultural characteristics that can influence their alignment with particular political parties. Ideally, stratification would also ensure race, gender, age, and other socioeconomic factors are also accounted for. However, these factors are almost impossible to determine while sending the survey. Therefore, the research team decided to ask demographic questions directly inside the survey. After data collection, the team will weigh data based on demographics. For example, if a certain county has a 30% black population (this statistic will be determined from the US census), but only 15% of the survey participants are black, then the pollsters may decide to count each black participant's responses twice.

Furthermore, the team will aim to sample approximately 1,000 people from each true swing state. This sample size is large enough to provide reliable conclusions but not so large that it resembles sampling with replacement. For true swing states and other close races, the team aims to survey 1,000 individuals. However, for non-swing states, if the final sample size is less than 1,000, it is not a significant issue.

@@

The team's ideal survey will be made using @qualtrics_qualtrics_nodate. It will attempt to ensure four things: no leading questions, no question order bias, no answer order bias, and the survey should identify non-engaged participants. To identify participants who are not engaged, the research team has decided to add a worthless question. This question is extremely simple and clearly has one correct answer. Therefore, if a participant selects the wrong answer, they are not engaged with the survey and their responses should be removed from the final data. An example of this is question number 7. The idea to use a worthless question was inspired by @pew_research_center_us_nodate.

@@

Answer bias, like order bias, occurs when the order of answer choices influences participants' selections. For instance, the first option can often be chosen more frequently. To prevent this, answer choices should be randomized upon survey entry. Questions 8 & 9 could benefit from this. Likewise, the team's hypothetical survey has not implemented this feature, but it is available in the @qualtrics_qualtrics_nodate paid plan.

@@

While the team's methodology and survey create a robust system, there are potential issues. The usage of weighting results based on a candidate's demographics can lead to error propagation. For instance, if a certain racial demographic population is only captured limitedly and that limited sample is far from representative, a few individuals in the population can have major impact on the polls prediction of what candidate will win the state. There is also a selection bias in terms of the monetary reward. Potentially, people who like monetary rewards would be more likely to engage in the survey and could therefore exhibit certain voting characteristics that create an unrepresentative sample.
